### 2022-07-20 draft.tex
Deploying to gh-pages from @ FrancisBanville/ms_maxent_networks@11ba21f9324a2b4bba7a74cddb24c7a98b456959 🚀

Shannon's entropy can only be calculated on conventional probability
distributions such as the joint degree distribution. This is an issue
when working with ecological networks, which are frequently represented
as adjacency matrices. For this reason, in this contribution, we also
use the SVD entropy as a measure of entropy, which is an application of
Shannon's entropy to the relative non-zero singular values of a
truncated singular value decomposition (t-SVD; Strydom, Dalla Riva, and
Poisot 2021) of a food web's Boolean adjacency matrix. We measured SVD
entropy as follows:

where \(s_i\) are the relative singular values of the adjacency matrix
(\(s_i = \sigma_i / \sum_{i = 1}^R \sigma_i\), where \(\sigma_i\) are
the singular values). Note that the distribution of relative singular
values is analogous to a probability distribution, with \(0 < s_i < 1\)
and \(\sum s_i = 1\). Thus, this measure also satisfies all four
properties above-mentioned, while being a proper measure of the internal
complexity of food webs (Strydom, Dalla Riva, and Poisot 2021).
Following Strydom, Dalla Riva, and Poisot (2021), we standardized this
measure with the rank \(R\) of the matrix (i.e.~\(J / \ln(R)\)) to
account for the difference in dimensions between networks (\emph{sensu}
Pielou's evenness; Pielou 1975). We will show how SVD entropy can be
used to predict the network of maximum entropy (i.e.~of maximum
complexity) heuristically given a predetermined set of constraints.

@@

In this section, we showed how important measures of food-web structure,
namely the degree distribution and the joint degree distribution, could
be derived with the principle of maximum entropy using minimal knowledge
on a biological community. This type of models, although useful to make
least-biased predictions on many network properties, can be hard to
apply for other measures. Indeed, there are dozens of measures of
network structure (Delmas et al. 2019) and many are not calculated with
mathematical equations, but with algorithms. Moreover, the applicability
of this method to empirical systems is limited by the state variables we
can actually measure and use. In the next section, we propose a more
flexible method to predict many measures of network structure
simultaneously, i.e.~by finding networks of maximum entropy
heuristically.

@@

@@

We define networks of maximum entropy as the configuration of the
adjacency matrix with the highest SVD entropy under a set of
constraints. As mentioned above, we used the SVD entropy as our measure
of entropy since it has been shown to be a reliable measure of food-web
complexity (Strydom, Dalla Riva, and Poisot 2021), in addition to having
the required properties of a proper measure of information entropy. We
thus seek to find the network with the highest complexity, or
randomness, that exactly reproduces specified constraints on its
structure. Our method is in contrast with maximum entropy graph models
that predict a probability distribution on networks under soft or hard
constraints (e.g., Park and Newman 2004; Cimini et al. 2019). We believe
our approach to be more flexible, easier to compute, while allowing
direct comparisons of empirical food webs with more complex networks
with similar structure.

@@

>>>>

@@

In this section, we showed how important measures of food-web structure,
namely the degree distribution and the joint degree distribution, could
be derived with the principle of maximum entropy using minimal knowledge
on a biological community. This type of models, although useful to make
least-biased predictions on many network properties, can be hard to
apply for other measures. Indeed, there are dozens of measures of
network structure (Delmas et al. 2019) and many are not calculated with
mathematical equations, but with algorithms. Moreover, the applicability
of this method to empirical systems is limited by the state variables we
can actually measure and use. In the next section, we propose a more
flexible method to predict many measures of network structure
simultaneously, i.e.~by finding heuristically the configuration of the
network having maximum entropy given partial knowledge of its emerging
structure.

@@

\hypertarget{from-shannons-to-svd-entropy}{%
\subsection{From Shannon's to SVD
entropy}\label{from-shannons-to-svd-entropy}}

The principle of maximum entropy can be applied on the network itself if
we decompose its adjacency matrix into a non-zero vector of relative
values. This is a necessary step when working with food webs, which are
frequently expressed as a matrix \(A = [a_{ij}]\) of Boolean values
representing the presence (\(a_{ij} = 1\)) or absence (\(a_{ij} = 0\))
of an interaction between two species \(i\) and \(j\). Knowing one or
many properties of a food web of interest (e.g., its number of species
and number of interactions), we can simulate its adjacency matrix
randomly by using these known ecological information to constrain the
space of potential networks. The entropy of this hypothetical matrix can
then be measured after decomposing it into appropriate values.
Simulating a series of networks until we find the one having the highest
entropy allows us to search for the most complex food-web configuration
given the ecological constraints used. This configuration is the least
biased one considering the information we had. In other words, the most
we can say about a network's adjacency matrix, without making more
assumptions than the ones given by our incomplete knowledge of its
structure, is the one of maximum entropy. Generating the most complex
network that corresponds to this structure allows us to explore more
easily other properties of food webs under MaxEnt.

Shannon's entropy can only be calculated on conventional probability
distributions such as the joint degree distribution. This is an issue
when working with ecological networks, which are represented as
adjacency matrices. For this reason, we need to use another measure of
entropy if we want to predict the network itself. We used the SVD
entropy as our measure of entropy, which is an application of Shannon's
entropy to the relative non-zero singular values of a truncated singular
value decomposition (t-SVD; Strydom, Dalla Riva, and Poisot 2021) of a
food web's Boolean adjacency matrix. We measured SVD entropy as follows:

where \(s_i\) are the relative singular values of the adjacency matrix
(\(s_i = \sigma_i / \sum_{i = 1}^R \sigma_i\), where \(\sigma_i\) are
the singular values). Note that the distribution of relative singular
values is analogous to a probability distribution, with \(0 < s_i < 1\)
and \(\sum s_i = 1\). Thus, this measure also satisfies all four
properties of an appropriate entropy measure (see introduction), while
being a proper measure of the internal complexity of food webs (Strydom,
Dalla Riva, and Poisot 2021). Following Strydom, Dalla Riva, and Poisot
(2021), we standardized this measure with the rank \(R\) of the matrix
(i.e.~\(J / \ln(R)\)) to account for the difference in dimensions
between networks (\emph{sensu} Pielou's evenness; Pielou 1975). We used
SVD entropy to predict the network configuration of maximum entropy
(i.e.~of maximum complexity) heuristically given a predetermined set of
constraints.

@@

@@

Our method is in contrast with maximum entropy graph models that predict
a probability distribution on networks under soft or hard constraints
(e.g., Park and Newman 2004; Cimini et al. 2019). We believe our
approach to be more flexible, easier to compute, while allowing direct
comparisons of empirical food webs with more complex networks with
similar structure.
